---
title: "Read Me: Replication Directory Overview"
subtitle: "Human Rights Violations in Space: Assessing the External Validity of Machine Geo-coded Vs. Human Geo-Coded Data"
date:  '`r format(Sys.time(), "%d %B, %Y")`'
output:
  rmarkdown::github_document:
  bookdown::pdf_document2:
    toc: no
    latex_engine: pdflatex
    number_sections: no
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(knitr)
library(kableExtra)

local_table <- function(data, c1_width = "4.5cm", c2_width = NULL, ...){
    if(knitr::is_latex_output()){
      data <- data.frame(lapply(data, function(x){gsub("`",   "", x)}),
                         stringsAsFactors = F)
      data <- data.frame(lapply(data, function(x){gsub("_",   "\\\\_", x)}),
                         stringsAsFactors = F)
      data <- data.frame(lapply(data, function(x){gsub("&",   "\\\\&", x)}),
                         stringsAsFactors = F)
      data <- data.frame(lapply(data, function(x){gsub("%",   "\\\\%", x)}),
                         stringsAsFactors = F)
      data <- data.frame(lapply(data, function(x){gsub("<br>", "\n", x)}), 
                         stringsAsFactors = F)  
      data <- data %>% mutate_all(linebreak)
      
      # Future note - for whatever reason, kable_styling() causes latex to 
      # freak out when passed a data frame containing cells with breaks
      
      if(is.null(c2_width)){
        data %>% kbl(format = "latex", booktabs = TRUE, escape = FALSE, ...) %>% 
          kable_styling(full_width = TRUE) %>% 
          column_spec(., column = 1, width = c1_width) %>% 
          row_spec(., row = 0, bold = TRUE) %>% 
          print
      } else{
        data %>% kbl(format = "latex", booktabs = TRUE, escape = FALSE, ...) %>% 
          column_spec(., column = 1, width = c1_width) %>% 
          column_spec(., column = 2, width = c2_width) %>% 
          row_spec(., row = 0, bold = TRUE) %>% 
          print
      }
    } else{
      data %>% 
        kbl(format = "simple", ...) %>% 
        print
  }
}

lnbrk <- function(){
  if(knitr::is_latex_output()){
    cat(paste("  \\"))
  } else{
    cat(paste("<br>"))
  }
}

```

  
```{r, results='asis'}
lnbrk()
```
 
# README OVERVIEW

```{r, results = "asis"}
dat <- bind_rows(
  c(a = "Author", b = "Logan Stundal, stund005@umn.edu"),
  c(a = "Date", b = "July 26, 2021"),
  c(a = "Article", b = "Human Rights Violations in Space: Assessing the External Validity of Machine Geo-coded Vs. Human Geo-coded Data"),
  c(a = "Journal", b = "Political Analysis"),
  c(a = "Co-Authors", b = "Bagozzi, Benjamin; Freeman, John; Holmes, Jennifer"),
  c(a = "Purpose", b = "This document explains script execution order to replicate all results in the paper / appendix as well as produce all figures and tables. Data imports and data exports are noted for each script.")
) %>% 
  local_table(data = ., c1_width = "2.5cm", col.names = NULL)
```


## Directory structure

* `Data/` - contains original data used as inputs for all models
    * `FARC-separate/` - contains separate Rdata files corresponding to period or annual data 
* `Scripts/` - contains all scripts to reproduce published results
* `Results/` - contains folders to store replication script outputs
    * `Replication-Estimates/` - stores exported estimated quantities
    * `Replication-Figures/` - stores exported figures
    * `Replication-Tables/` - stores exported tables
    * `Published-Models` - stores published models as compressed `.Rdata` files

## Replication files are located at:  
  - [GitHub repository](https://github.com/loganstundal/EventData-Space-Colombia/tree/main/Replication)  
  - [Harvard Dataverse]()

\pagebreak



# Hardware and Software
All models and analysis were conducted using `r R.version.string` running on a PC using Windows 10 x64 (build 19043) with an Intel Core i5-6200U CPU (2.30GHz) and 8Gb of DDR3 memory (1867 MHz).

The primary analysis in this paper was conducted using R-INLA version 21.02.23 compiled on Feb 22, 2021. The tar.gz installation file for this version of INLA is provided here:  
  - https://inla.r-inla-download.org/R/stable/src/contrib/
  
  - https://inla.r-inla-download.org/R/stable/src/contrib/INLA_21.02.23.tar.gz

The following packages are called in scripts 1-7. Code to install these packages is located in `Scripts/0-master.R`.

```{r, results = "asis"}
dat <- data.frame(
  Package = c("cowplot", "dplyr", "forcats", "ggplot2", "ggrepel", "kableExtra", "ProbitSpatial", "pROC", "purrr", "raster", "sandwich", "scales", "sf", "spdep", "stringr", "tibble", "tidyr"),
  Version = c("1.1.1", "1.0.7", "0.5.1", "3.3.3", "0.9.1", "1.3.4", "1.0", "1.17.0.1", "0.3.4", "3.4.5", "3.0.0", "1.1.1", "0.9.7", "1.1.5", "1.4.0", "3.0.6", "1.1.3")
) %>% 
  local_table(data = ., c1_width = "2in", c2_width = "1in")
```

\pagebreak



# Data

All data for this project are stored in the `Data/` subdirectory which contains three files:

* **`/farc_events.Rdata`** contains all data necessary to replicate the analysis. This file contains an r-object called `dat` which is a named list comprising of 4 data frames: `2002-2004`, `2005-2007`, `2008-2009`, and `2002-2009` each corresponding to a temporal aggregation of the data implied by the name. `FARC-separate/` contains these data frames as separate `Rdata` files. These data frames all contain the following variables:

```{r, results = "asis"}
dat <- bind_rows(
  c(a = "department",   b = "Colombia Admin. level-1, Department name"),
  c(a = "municipality", b = "Colombia Admin. level-2, Municipality name"),
  c(a = "yr_grp",       b = "Year periodization"),
  c(a = "cinep",        b = "FARC activity, count (CINEP source)"),
  c(a = "cinep_bin",    b = "FARC activity, binary-indicator (CINEP source)"),
  c(a = "icews",        b = "FARC activity, count (ICEWS source)"),
  c(a = "icews_bin",    b = "FARC activity, binary-indicator (ICEWS source)"),
  c(a = "icews_cinep_under", b = "FARC activity, ICEWS underreporting relative to CINEP, binary-indicator"),
  c(a = "ged",          b = "FARC activity, count (GED source)"),
  c(a = "ged_bin",      b = "FARC activity, binary-indicator (GED source)"),
  c(a = "ged_cinep_under", b = "FARC activity, GED underreporting relative to CINEP, binary-indicator"),
  c(a = "area_km2", b = "Municipality area"),
  c(a = "area_km2_ln", b = "Municipality area, logged"),
  c(a = "distance_bogota_km_ln", b = "Distance (km) between municipality centroid and Bogota, logged"),
  c(a = "distance_bogota_km", b = "Distance (km) between municipality centroid and Bogota"),
  c(a = "pop_sum", b = "Population count at municipality level"),
  c(a = "pop_sum_ln", b = "Population count at municipality level, logged"),
  c(a = "terrain_ri_mean_m", b = "Terrain roughness indicator, municipality average (meters)"),
  c(a = "centroid_mun_lat", b = "Municipality centroid, latitude"),
  c(a = "centroid_mun_long", b = "Municipality centroid, longitude"),
) %>% 
  local_table(data = ., col.names = c("Variable name", "Description"))
```

* **`colombia.Rdata`** - an unprojected simple feature of Colombia's international border used to provide a sharply contrasting boundary in maps containing posterior mean, probability, and standard deviation estimates of the Gaussian random field. 
* **`colombia2.Rdata`** - an unprojected simple feature object of Colombia's level-2 administrative boundaries (Municipalities) used for mapping purposes.

\pagebreak



# Replication results

The R-scripts stored in the `Scripts/` directory generate outputs to produce all figures or tables presented in the main article or appendix. These exports are sored in one of three sub-directories in the `Results/` folder:

  - **`Replication-Estimates/`** - contains `.Rdata` files with model estimates used to reproduce all tables and figures
  - **`Replication-Figures/`** - contains `.png` files of all figure exports
  - **`Replication-Tables/`** - contains outputs used to create all tables in main article and appendix. 
  - **`Published-Models/`** - contains compressed `.Rdata` files with pre-estimated models from executing code in `2-models-spde.R` or `3-models-spem.R` 
  
R-INLA models with random fields have large file sizes (~130Mb * 5 models * 4 time cuts = ~2.5Gb). To reduce file sizes, after estimating models in the primary `2-models-spde.R` script,  quantities of interest necessary for figures or tables are extracted or computed and exported as vectors rather than complete model objects. The compressed full model results are stored as an Rdata object in `Published-Models/`

# Scripts

**Note** - All scripts assume the top-level `Replication/` folder is set as the working directory.

The following tables provide an overview of each script in the `Scripts/` directory. Each table indicates what data sources are imported - either raw data from `Data/` or estimated quantities from `Replication-Estimates/`. Each table also indicates what component of the paper the script exports (either a table or figure) as well as where that element is stored within the Replication directory.

```{r, results = 'asis'}
# Template for script tables
dat <- bind_rows(
  c(a = "Purpose",      b = "This top-level script contains code to install all packages used in scripts 1-7 as well as code that automates the execution of scripts 1-7."),
  c(a = "Runtime",      b = "42.6 minutes - total run time for all scripts")
) %>% 
  local_table(data = ., 
              c1_width  = "1.1in", 
              c2_width  = "5.1in",
              col.names = c("Script Name", "0-master.R"))
```
  
```{r, results='asis'}
lnbrk()
```
  
```{r, results = 'asis'}
dat <- bind_rows(
  c(a = "Purpose",      b = "Reproduces the descriptive data overview and comparison presented in main paper Section 3.2 including Figures 1 & 2 and Tables 1 & 2. This script also produces Appendix Table 1.<br>"),
  c(a = "Imports",      b = "`Data/farc_events.Rdata`<br>`Data/colombia.Rdata`<br>`Data/colombia2.Rdata`<br>"),
  c(a = "Exports",      b = "`Replication-Figures/figure_main_1.png` - Figure 1, observed FARC Events map<br>`Replication-Figures/figure_main_2.png` - Figure 2, selected remote and non-remote municipalities map<br>`Replication-Tables/table_main_1.txt` - Table 1, full cross-section confusion matrix<br>`Replication-Tables/table_main_2.txt` - Table 2, selected remote vs. non-remote municipalities confusion matrices<br>`Replication-Tables/table_appendix_1.txt` - Table A1, selected remote and non-remote municipalities table<br>"),
  c(a = "Dependencies", b = "`dplyr`, `tidyr`, `forcats`, `stringr`, `ggplot2`, `ggrepel`, `sf`, `purrr`, `kableExtra`"),
  c(a = "Runtime",      b = "20 seconds")
) %>% 
  local_table(data = ., 
              c1_width  = "1.1in", 
              c2_width  = "5.1in",
              col.names = c("Script Name", "1-descriptive.R"))
```
  
```{r, results='asis'}
lnbrk()
```

```{r, results = 'asis'}
dat <- bind_rows(
  c(a = "Purpose",      b = "Reproduces INLA models with SPDEs reported graphically in the main draft as well as in table format in the Appendix.<br>"),
  c(a = "Imports",      b = "`Data/farc_events.Rdata`<br>"),
  c(a = "Exports",      b = "`Replication-Estimates/parameter-data.Rdata` - All model parameter and HPD estimates<br>`Replication-Estimates/field-data.Rdata` - Projected estimates of Gaussian field mean and SD<br>`Replication-Estimates/range-data.Rdata` - Spatial field decay and range estimates<br>`Replication-Estimates/pred-data.Rdata` - Model predicted outcomes on probability scale<br>`Published-Models/published-models-spde.Rdata` - Compressed INLA model estimates presented in the published article<br>"),
  c(a = "Dependencies", b = "`dplyr`, `tibble`, `INLA`"),
  c(a = "Runtime",      b = "27 minutes")
) %>% 
  local_table(data = ., 
              c1_width  = "1.1in", 
              c2_width  = "5.1in",
              col.names = c("Script Name", "2-models-spde.R"))
```
  
```{r, results='asis'}
lnbrk()
```

```{r, results = 'asis'}
dat <- bind_rows(
  c(a = "Purpose",      b = "Reproduces discrete spatial probit error models (spem) presented in Appendix Tables A3 and A4 as well as Figure A3 (spem ROC comparison) and Figure A4 (spem predicted probabilities). This script and associated export also provides estimates for non-spatial probits presented in the Appendix.<br>"),
  c(a = "Imports",      b = "`Data/farc_events.Rdata`<br>"),
  c(a = "Exports",      b = "`Published-Models/published-models-spem.Rdata` - Estimated SPEM and non-spatial probit models<br>`Replication-Figures/figure_appendix_4.png` - Figure A4, SPEM mapped predicted outcome probabilities<br>`Replication-Tables/table_appendix_3.txt` - Table A3, Probit (non-spatial) 2002-2009<br>`Replication-Tables/table_appendix_4.txt` - Table A4, SPEM 2002-2009<br>"),
  c(a = "Dependencies", b = "`dplyr`, `tidyr`, `tibble`, `ggplot2`, `stringr`, `purrr`, `forcats`, `sf`, `ProbitSpatial`, `sandwich`, `spdep`, `kableExtra`"),
  c(a = "Runtime",      b = "13 minutes")
) %>% 
  local_table(data = ., 
              c1_width  = "1.1in", 
              c2_width  = "5.1in",
              col.names = c("Script Name", "3-models-spem.R"))
```
  
```{r, results='asis'}
lnbrk()
```

```{r, results = 'asis'}
# Template for script tables
dat <- bind_rows(
  c(a = "Purpose",      b = "Reproduces Figures 3 and 4 from the main paper which present INLA model parameter estimates and credibility intervals for included regressors as well as GRMF parameters.<br>"),
  c(a = "Imports",      b = "`Replication-Estimates/parameter-data.Rdata`<br>"),
  c(a = "Exports",      b = "`Replication-Figures/figure_main_3.png` - Figure 3, Observed FARC SPDE model estimates<br>`Replication-Figures/figure_main_4.png` - Figure 4, Underreporting FARC SPDE model estimates<br>"),
  c(a = "Dependencies", b = "`dplyr`, `tidyr`, `stringr`, `forcats`, `ggplot2`, `cowplot`"),
  c(a = "Runtime",      b = "6 seconds")
) %>% 
  local_table(data = ., 
              c1_width  = "1.1in", 
              c2_width  = "5.1in",
              col.names = c("Script Name", "4-figures\\_coefplots.R"))
```
  
```{r, results='asis'}
lnbrk()
```

```{r, results = 'asis'}
# Template for script tables
dat <- bind_rows(
  c(a = "Purpose",      b = "Produces figure 5 which presents ROCs and AUC estimates comparing ICEWS and GED model performance against CINEP ground truth observations.<br>"),
  c(a = "Imports",      b = "`Data/farc_events.Rdata`<br>`Replication-Estimates/pred-data.Rdata`<br>`Published-Models/published-models-spem.Rdata`<br>"),
  c(a = "Exports",      b = "`Replication-Figures/figure_main_5.png` - ROC SPDE (Observed FARC)<br>`Replication-Figures/figure_appendix_3.png` - Figure A3, ROC SPEM (Observed FARC)<br>`Replication-Figures/figure_appendix_5.png` - Figure A4, ROC SPEM (Underreporting FARC)<br>`Replication-Figures/figure_appendix_6.png` - Figure A6, ROC SPDE (Underreporting FARC)<br>`Replication-Figures/figure_appendix_7.png` - Figure A7, ROC SPDE (Observed and Underreporting 2008-2009 period)<br>"),
  c(a = "Dependencies", b = "`dplyr`, `tidyr`, `ggplot2`, `stringr`, `purrr`, `pROC`, `cowplot`"),
  c(a = "Runtime",      b = "33 seconds")
) %>% 
  local_table(data = ., 
              c1_width  = "1.1in", 
              c2_width  = "5.1in",
              col.names = c("Script Name", "5-figures-ROCs.R"))
```
  
```{r, results='asis'}
lnbrk()
```

```{r, results = 'asis'}
# Template for script tables
dat <- bind_rows(
  c(a = "Purpose",      b = "Produces Appendix Tables A5-A8 which present INLA SPDE model estimates in table format with median and 95% HPD estimates.<br>"),
  c(a = "Imports",      b = "`parameter-data.Rdata`<br>"),
  c(a = "Exports",      b = "`Replication-Tables/table_appendix_5.txt` - Table A5, SPDE 2002-2009<br>`Replication-Tables/table_appendix_6.txt` - Table A6, SPDE 2002-2004<br>`Replication-Tables/table_appendix_7.txt` - Table A7, SPDE 2005-2007<br>`Replication-Tables/table_appendix_8.txt` - Table A8, SPDE 2008-2009<br>"),
  c(a = "Dependencies", b = "`dplyr`, `tidyr`, `kableExtra`"),
  c(a = "Runtime",      b = "2 seconds")
) %>% 
  local_table(data = ., 
              c1_width  = "1.1in", 
              c2_width  = "5.1in",
              col.names = c("Script Name", "6-tables.R"))
```
  
```{r, results='asis'}
lnbrk()
```

```{r, results = 'asis'}
# Template for script tables
dat <- bind_rows(
  c(a = "Purpose",      b = "Produces figures related to posterior Gaussian Markov Random Field. Figure 8 - posterior field maps and Figure 9 - Posterior spatial error range and decay.<br>"),
  c(a = "Imports",      b = "`Replication-Estimates/field-data.Rdata`<br>`Replication-Estimates/range-data.Rdata`<br>`Data/colombia.Rdata`<br>"),
  c(a = "Exports",      b = "`Replication-Figures/figure_appendix_8.png` - Figure 8, SPDE 2002-2009 model GMRF estimate maps<br>`Replication-Figures/figure_appendix_9.png` - Figure 9, SPDE 2002-2009 model GMRF error correlation range and decay<br>"),
  c(a = "Dependencies", b = "`INLA`, `sf`, `dplyr`, `tidyr`, `ggplot2`, `purrr`, `raster`, `cowplot`, `scales`"),
  c(a = "Runtime",      b = "1.2 minutes")
) %>% 
  local_table(data = ., 
              c1_width  = "1.1in", 
              c2_width  = "5.1in",
              col.names = c("Script Name", "7-figures-field\\_range-estimates.R"))
```


<!-- rmarkdown::render("README.rmd", "pdf_document")
rmarkdown::render("README.rmd", "html_document") -->


```{r, results = 'asis', eval =  FALSE}
# Template for script tables
dat <- bind_rows(
  c(a = "Purpose",      b = ""),
  c(a = "Imports",      b = ""),
  c(a = "Exports",      b = ""),
  c(a = "Dependencies", b = ""),
  c(a = "Runtime",      b = "")
) %>% 
  local_table(data = ., 
              c1_width  = "1.1in", 
              c2_width  = "5.1in",
              col.names = c("Script Name", "THE SCRIPT.R"))
```

